---
title: "Interactive Analysis of Formula 1 \nQualifying Data"
author:
  - name: Norah Kuduk
  - name: Matt Higham (Advisor)
affiliation:
    address: Department of Statistics, St. Lawrence University
column_numbers: 3
poster_height:	"36in"	
poster_width:	"48in"
primary_colour: "#aa1a0a"
secondary_colour: "#aa1a0a"
titlebox_bordercol:	"#aa1a0a"
titlebox_borderwidth:	"0.5cm"
font_family:	"Calibri"
logoright_name: "figures/slu_white.png"
logoleft_name: "figures/f1-logo.png"
sectitle_bgcol:	"#aa1a0a"
author_textcol:	"#ffffff"
sectitle2_textcol:	"#aa1a0a"
link_col:	"#aa1a0a"
output: 
  posterdown::posterdown_html:
    self_contained: false
    

bibliography: packages.bib
---

```{r setup, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
library(tidyverse)
library(f1dataR)
library(f1animateR)
library(knitr)
library(ggrepel)
```

# What is Formula 1 (F1)?

- F1 is an international motorsport series governed by the Fédération Internationale de l'Automobile (FIA), involving 20 drivers across 10 teams that compete in 18-24 race weekends across a season.

- Each weekend features practice, qualifying, and race sessions, with qualifying laps representing peak performance used to decide the starting order for the race. 

- A car’s fastest lap from a qualifying session highlights driver skill, car setup efficiency, and team strategy under peak performance conditions. 

# Objectives

1. Compare driver performance across races/seasons, thinking about historically important or interesting sessions. 

2. Provide an interactive, visual platform to compare qualifying laps for users.
   

# Methods & Tools

- Using `f1dataR` package (R version of `fastf1` for Python, Ergast API, etc.).

- this is an extra bullet point to tray and fix spacing

- Creation of a local R package (`f1animateR`) that uses some functions from `f1dataR`, as well as helper functions when using `gganimate`.

- Using Shiny to create the app, hosted on the St. Lawrence University server.

# Results

- As a case study, we are looking at laps from the São Paulo Grand Prix, both across seasons and across different drivers.

## Variable Details
| Variable Name | Description        |
|:----:|:-----------------|
| Driver   | A 3-letter unique abbriviation of a drivers last name.      |
| Season   | The year of the session.        |
|    |                     |
| Lap Time    | The time (in seconds) of the lap.        |
| Compound    | The name of the tire compund the driver is using. Different compounds have different grip levels.  |

## Season Comparison

- The season comparison highlights how wet weather conditions in 2023 led to significantly slower lap times compared to the dry conditions in 2024, as the positions between the two seasons are drastically far from each other. 

```{r tiredegplot, out.width='100%', fig.align='center', fig.cap='Plot of the five fastest lap times and tire degredation across the two seasons.', fig.height=3}
fetch_quali_data_seasons <- function(season_a, season_b, round) {
  tryCatch(
    {
      results_a <- load_session_laps(season_a, round, "Q") |> mutate(season = season_a)
      results_b <- load_session_laps(season_b, round, "Q") |> mutate(season = season_b)
      if (is.null(results_a) || is.null(results_b) || nrow(results_a) == 0 || nrow(results_b) == 0) {
        return(NULL)
      } else {
        results <- bind_rows(results_a, results_b)
        return(results)
      }
      
    },
    error = function(e) NULL
  )
}

quali_laps <- fetch_quali_data_seasons(2024, 2023, "São Paulo Grand Prix") |>
  filter(driver == "NOR") |>
  filter(lap_time > 0) |>
  select(driver, season, lap_time, lap_number, compound) |>
  arrange(lap_time) |>
  group_by(season) |>
  slice(1:5) |>
  ungroup() |>
  arrange(desc(lap_time)) |>
  mutate(index = row_number()) |> 
  mutate(season = as.factor(season))

ggplot(quali_laps, aes(x = index, y = lap_time)) +
  geom_segment(aes(x = index, xend = index, y = 0, yend = lap_time, color = compound, linetype = season)) +
  geom_point(size = 1.5, aes(color = compound), alpha = 0.7) +
  scale_color_manual(values = c("SOFT" = "red", "MEDIUM" = "gold", "HARD" = "white", "WET" = "blue", "INTERMEDIATE" = "#29AF7FFF",
                                "HYPERSOFT" = "lightpink", "ULTRASOFT" = "purple", "SUPERSOFT" = "tomato", "SUPERHARD" = "orange")) +
  coord_flip() +
  labs(
    title = "Tire Performance Degradation for Lando Norris",
    y = "Lap Time (seconds)",
    color = "Tire Compound",
    linetype = "Season"
  ) +
  theme_minimal() +
  theme(legend.position = "top", legend.text = element_text(color = "#1C1C1C"),
        legend.title = element_text(color = "#1C1C1C", face = "bold"),
        legend.box.background = element_rect(color="white"),
        axis.title.y = element_blank(), axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        grid.major = element_blank(), grid.minor = element_blank())
```
```{r resultstable, out.width='100%'}
fetch_quali_data <- function(season, round) {
  tryCatch(
    {
      results <- load_session_laps(season, round, "Q")
      if (is.null(results) || nrow(results) == 0) return(NULL)
      results
    },
    error = function(e) NULL
  )
}

year_1 <- get_session_drivers_and_teams(2024, "São Paulo Grand Prix", "Q") |>
  mutate(driver_code = abbreviation) |>
  select(-abbreviation)

year_2 <- get_session_drivers_and_teams(2023, "São Paulo Grand Prix", "Q") |>
  mutate(driver_code = abbreviation) |>
  select(-abbreviation)

data <- bind_rows(year_1, year_2) |> filter(driver_code == "NOR")

quali_laps_a <- fetch_quali_data(2024, "São Paulo Grand Prix") |>
  filter(driver == "NOR") |>
  filter(lap_time > 0) |>
  mutate(season = 2024)

quali_laps_b <- fetch_quali_data(2023, "São Paulo Grand Prix") |>
  filter(driver == "NOR") |>
  filter(lap_time > 0) |>
  select(driver, lap_time, lap_number, compound) |>
  mutate(season = 2023)

data <- bind_rows(quali_laps_a, quali_laps_b) |>
  unique() |>
  arrange(lap_time) |>
  group_by(season) |>
  filter(lap_time == min(lap_time)) |>
  select(driver, season, lap_time, compound) 

kable(data,caption = 'Qualifying results for Lando Norris.', align = 'c',"html")
```

```{r animsampleplot, out.width='100%', echo=FALSE, warnings=FALSE, fig.cap='Snapshot of a late point in the lap where cars are very far apart (about 69 seconds in).', fig.height=5}
brazil24 <- get_quali_telemetry(laps = "fastest", season = 2024, round = 21, driver = "NOR", verbose = FALSE)
brazil23 <- get_quali_telemetry(laps = "fastest", season = 2023, round = 20, driver = "NOR", verbose = FALSE)

# smooth Abu Dhabi track data
brazil_track <- smooth_track(brazil24)
start_coord <- brazil_track |> slice(1)

# det team and driver colors
color_24 <- driver_color("NOR")
color_23 <- "darkorange4"

# Add driver labels and combine data
brazil24 <- brazil24 |> mutate(season = 2024)
brazil23 <- brazil23 |> mutate(season = 2023)

plot_data <- bind_rows(brazil24, brazil23) |> mutate(season = as.factor(season)) 

time_data <- plot_data |> group_by(season) |> filter(time > 68.9) |> filter(time < 68.98)

# Pick two consecutive points near the start of the track to show direction
arrow_start <- brazil24 |> slice(3)
arrow_end <- brazil24 |> slice(10)

 # Static plot
static_plot <- ggplot() +
  geom_path(data = brazil_track, aes(x = x2, y = y2, group = 1),
            linewidth = 4, color = "white") +
  geom_point(data = start_coord, aes(x = x2, y = y2),
             color = "black", shape = 18, size = 3) +
  geom_point(data = time_data, aes(x = x, y = y, group = season, color = season),
             size = 2) +
  geom_segment(
    data = NULL,
    aes(x = arrow_start$x, y = arrow_start$y,
        xend = arrow_end$x, yend = arrow_end$y),
    arrow = arrow(length = unit(0.1, "cm"), type = "closed"),
    color = "gray40",
    linewidth = 0.7
  ) +
  scale_color_manual(values = c("2024" = color_24, "2023" = color_23)) +
  theme_track() +
  labs(title = "2024 vs. 2023 Norris Qualifying Lap",
     subtitle = "São Paulo Grand Prix",
     x = NULL, y = NULL)

# Correct the track alignment
corrected_plot <- corrected_track(static_plot, brazil24) + geom_segment(
    data = NULL,
    aes(x = arrow_start$x, y = arrow_start$y,
        xend = arrow_end$x, yend = arrow_end$y),
    arrow = arrow(length = unit(0.1, "cm"), type = "closed"),
    color = "gray40",
    linewidth = 0.7
  ) 

corrected_plot
```

## Driver Comparison
- The driver comparison shows how in constant conditions, differences in drivers/teams can have drastic effect on qualifying times, but even between 1st and 20th, the margins are very small. 

```{r drivercomp, out.width='100%'}
quali_laps <- fetch_quali_data(2023, "São Paulo Grand Prix") |>
  filter(driver == "VER" | driver == "ZHO") |>
  filter(lap_time > 0) |>
  select(driver, lap_time, lap_number, compound) |>
  mutate(season = 2023) |> 
  group_by(driver) |>
  filter(lap_time == min(lap_time)) |>
  select(driver, season, lap_time, compound) 

kable(quali_laps, caption = 'Qualifying results for Verstappen and Zhou', align = 'c',"html")
```

```{r driveranimplot, out.width='100%', fig.align='center', fig.cap='Snapshot of a late point in the lap where the two drivers\' cars are far apart (about 63 seconds in).'}
ver <- get_quali_telemetry(laps = "fastest", season = 2023, round = 20, driver = "VER", verbose = FALSE)
zho <- get_quali_telemetry(laps = "fastest", season = 2023, round = 20, driver = "ZHO", verbose = FALSE)

# smooth Abu Dhabi track data
brazil_track <- smooth_track(ver)
start_coord <- brazil_track |> slice(1)

# det team and driver colors
ver_color <- constructor_color("redbull")
zho_color <- constructor_color("sauber")

# Add driver labels and combine data
ver <- ver |> mutate(driver = "VER")
zho <- zho |> mutate(driver = "ZHO")

plot_data <- bind_rows(ver, zho) |> mutate(driver = as.factor(driver))

time_data <- plot_data |> group_by(driver) |> filter(time > 63.2) |> filter(time < 63.4)

# Pick two consecutive points near the start of the track to show direction
arrow_start <- ver |> slice(3)
arrow_end <- ver |> slice(10)

# Static plot
static_plot <- ggplot() +
  geom_path(data = brazil_track, aes(x = x2, y = y2, group = 1),
            linewidth = 4, color = "white") +
  geom_point(data = start_coord, aes(x = x2, y = y2),
             color = "black", shape = 18, size = 3) +
  geom_point(data = time_data, aes(x = x, y = y, group = driver, color = driver),
             size = 2) +
  scale_color_manual(values = c("VER" = ver_color, "ZHO" = zho_color)) +
  geom_segment(
    data = NULL,
    aes(x = arrow_start$x, y = arrow_start$y,
        xend = arrow_end$x, yend = arrow_end$y),
    arrow = arrow(length = unit(0.1, "cm"), type = "closed"),
    color = "gray40",
    linewidth = 0.7
  ) +
  theme_track() +
  labs(title = "Verstappen vs. Zhou Qualifying Lap",
       subtitle = "São Paulo Grand Prix 2023",
       x = NULL, y = NULL)

# Correct the track alignment
corrected_plot <- corrected_track(static_plot, ver)

corrected_plot
```

# Conclusion

Data is a very large part of Formula 1. Visual analytics offer deeper insight into qualifying performance for the average viewer. 
Interactive tools like Shiny make F1 data more accessible and fun.

```{r, include=FALSE}
knitr::write_bib(c('knitr','rmarkdown','posterdown','pagedown'), 'packages.bib')
```

# References

`f1dataR`: Formula 1 data access package. Available at: [https://cran.r-project.org/package=f1dataR](https://cran.r-project.org/package=f1dataR)

What is Formula 1? Available at: [https://www.formula1.com/en/page/what-is-f1](https://www.formula1.com/en/page/what-is-f1)
