# Playing with Interesting Qualifying Laps

## Loading Packages
```{r}
library(f1dataR)
library(tidyverse)
library(dplyr)
library(gganimate)
library(gifski)

# run util-functions
```

## Full Example of Using Functions

### Lando Norris Dutch GP 2024
```{r}
# Get the data
nor_dutch_quali_2024 <- load_driver_telemetry(season = 2024, round = 15, driver = "NOR", session = "Q", laps = "fastest")

dutch_2024_track <- smooth_track(nor_dutch_quali_2024)
```

```{r}
mcl_color <- constructor_color("mclaren")
```

```{r}
uncorrected <- ggplot() +
  geom_path(data = dutch_2024_track, aes(x=x2, y=y2, group=1), linewidth=8, color = "white") +
  geom_point(data = nor_dutch_quali_2024, aes(x=x, y=y, group = NA),
             alpha = 0.7, color = mcl_color, size=3)+
  labs(x=NULL, y=NULL) +
  theme_track() +
  labs(title = "Norris Qualifying Lap", 
  subtitle = "Dutch Grand Prix 2024")
```

```{r}
p_nor_dutch_2024 <- corrected_track(uncorrected, nor_dutch_quali_2024)
```

```{r}
anim <- p_nor_dutch_2024 +
  transition_reveal(along = time)+
  ease_aes('linear') +
  labs(caption = "Time: {sprintf('%.3f', frame_along)} s
       Speed:  {nor_dutch_quali_2024$speed[which.min(abs(nor_dutch_quali_2024$time-frame_along))]} kph")
```

```{r}
animate(anim, width = 700, height = 700, fps = 10)
```

## TODO

speed smoothing -> every 4 frames take the average? can test this

start/finish line representation -> strip at start/finish coordinates

## Further Exploration

### Silverstone 2022 vs Silverstone 2023

Silverstone 2022 had a completely wet qualifying, meaning drivers were on inters for Q1, Q2, and Q3. Would be interesting to explore average lap times, as well as look at a plot of fastest wet quali lap vs dry quali lap. Could either compare pole position laps, or could compare drivers laps, or both

Carlos Sainz Pole 2022 - 1:40.983
Max Verstappen 2nd 2022 - 1:41.055 (*Had fastest overall lap in qualifying, but happened in Q2)

Max Verstappen Pole 2023 - 1:26.720
Carlos Sainz 5th 2023 - 1:27.148

### United States 2024x

Most recent race, Q3 interupted by a yellow flag in the last seconds of running, meant that Lando Norris's lap from the first runs of qualifying was good enough for pole. Would be interesting to track data for Verstappen's final lap up until the yellow flag to see if he was considerably ahead of Norris before the yellow flag. 

**Have to figure out how to pull telemetry data from a lap that is not fastest

### Monza 2019

Couple of options here, both are good to look at for interesting qualifying, and not just fastest laps

Norris gave a tow to Sainz during Q2, knowing he would be starting from the back of the grid. Does a tow really work? Can you tell visually or in the data where it has the most affect?

Q3 also interesting, only Leclerc and Sainz were able to have second flying laps, interesting to see how much the track develops over the course of a qualifying session.

